Incremental λ-Calculus in Cache-Transfer Style

### 2017-07-11 popl18/source-language.tex
Introduce syntax for tuples
More grammar

Our syntax for base terms~$\sterm$ represents $\lambda$-lifted
programs in A'-normal form (\cref{sec:motivating-example}).
A term can be either a bound variable~$\tx$, or a {\bf let}
binding of $\ty$ in subterm~$\sterm$ to either a new
tuple~$\stuple{\many\tx}$ ($\slet{\ty = \stuple{\many\tx}}\sterm$), or
the result of calling function~$\tf$ on argument~$\tx$
($\slet{\ty = \sapp{\tf}{\tx}}{\sterm}$).  Both~$\tf$ and~$\tx$ are
variables to be looked up in the environment.  Terms cannot contain
$\lambda$-abstractions as they have been $\lambda$-lifted to top-level
definitions, which we encode as closures in the initial environments.
%
Unlike standard ANF we add no special syntax for function
calls in tail position (see \cref{sec:tail-calls} for a discussion
about this limitation).
%
We often need to inspect the result of
a function call~$\tf\,\tx$, which is not a valid term in our syntax.
To alleviate this, we write~``$\scall\tf\tx$''
for ``$\slet{\ty = \tf\;\tx}\ty$'' where $\ty$ is
chosen fresh.

@@

Nil and replacement changes can be used to update both
constants, tuples, primitives and closures, while tuple change can
only update tuples, literal change can only update literals and
closure changes can only update closures. Indeed, $\oplus$ on literals
is defined \textit{via} some interpretation function~$\delta_{\oplus}$. A
replacement change overrides the current value~$\sclosedvalue$ with a
new one~$\sclosedvalue'$.
Change application for a closure
ignores~$\idterm$ instead of combining it with $\sterm$ somehow. This may seem
surprising, but we only need $\oplus$ to behave well for valid changes (as shown
by \cref{lemma:crel-oplus}): for valid closure changes,
$\idterm$ must behave similarly to $\iderive\sterm$ anyway,\pg{Not formal, not
  visible in the definition, but true for a suitable meaning of ``similarly'';
  but I'm clearly not claiming a theorem, just giving a hint for clarity.}
so only environment updates matter. This definition also avoids having to modify
terms at runtime, which would have been difficult to implement safely.
\pg{$\oplus$ is not used by derivatives, so commenting this part out
A change application for a closure
ignores~$\idterm$, which may seems strange at first but is needed to
but is needed to
get guarantees about the evaluation of derivatives.}
% With this restriction, we avoid the case of a new
% closure code that would refer to a variable undefined in the closure
% environment.
% \pg{I can imagine fixes to that problem, so I'd propose to stick to the
% stronger arguments.}

>>>>

Our syntax for base terms~$\sterm$ represents $\lambda$-lifted
programs in A'-normal form (\cref{sec:motivating-example}).
We write $\many\tx$ for a sequence of variables of some unspecified length
$\tx_1, \tx_2, \ldots, \tx_m$.
A term can be either a bound variable~$\tx$, or a {\bf let}
binding of $\ty$ in subterm~$\sterm$ to either a new
tuple~$\stuple{\many\tx}$ ($\slet{\ty = \stuple{\many\tx}}\sterm$), or
the result of calling function~$\tf$ on argument~$\tx$
($\slet{\ty = \sapp{\tf}{\tx}}{\sterm}$).  Both~$\tf$ and~$\tx$ are
variables to be looked up in the environment.  Terms cannot contain
$\lambda$-abstractions as they have been $\lambda$-lifted to top-level
definitions, which we encode as closures in the initial environments.
%
Unlike standard ANF we add no special syntax for function
calls in tail position (see \cref{sec:tail-calls} for a discussion
about this limitation).
%
We often need to inspect the result of
a function call~$\tf\,\tx$, which is not a valid term in our syntax.
To alleviate this, we write~``$\scall\tf\tx$''
for ``$\slet{\ty = \tf\;\tx}\ty$'' where $\ty$ is
chosen fresh.

@@

Nil and replacement changes can be used to update
constants, tuples, primitives and closures, while tuple change can
only update tuples, literal change can only update literals and
closure changes can only update closures. Indeed, $\oplus$ on literals
is defined \textit{via} some interpretation function~$\delta_{\oplus}$. A
replacement change overrides the current value~$\sclosedvalue$ with a
new one~$\sclosedvalue'$.
Change application for a closure
ignores~$\idterm$ instead of combining it with $\sterm$ somehow. This may seem
surprising, but we only need $\oplus$ to behave well for valid changes (as shown
by \cref{lemma:crel-oplus}): for valid closure changes,
$\idterm$ must behave similarly to $\iderive\sterm$ anyway,\pg{Not formal, not
  visible in the definition, but true for a suitable meaning of ``similarly'';
  but I'm clearly not claiming a theorem, just giving a hint for clarity.}
so only environment updates matter. This definition also avoids having to modify
terms at runtime, which would have been difficult to implement safely.
\pg{$\oplus$ is not used by derivatives, so commenting this part out
A change application for a closure
ignores~$\idterm$, which may seems strange at first but is needed to
but is needed to
get guarantees about the evaluation of derivatives.}
% With this restriction, we avoid the case of a new
% closure code that would refer to a variable undefined in the closure
% environment.
% \pg{I can imagine fixes to that problem, so I'd propose to stick to the
% stronger arguments.}
